---
title: "RI Community Studios Impact Report Season 1"
edition: "Season 1 (April to June 2024)"
format:
  PrettyPDF-typst:
    typst-logo: 
      path: "./communitystudioslogo.png"
    mainfont: "Arial"
    echo: false
    warning: false
    message: false
    fig-width: 7
    margin:
      x: 1.5cm
      y: 1.5cm
  html: 
    echo: false
    warning: false
    message: false
params:
  sheet: ""
editor: source
---

DaSL Community Studios (CommStudios) are events designed to catalyze community and conversations around shared workflows and interests in biomedical data science topics. Specifically, in the Research informatics CommStudios, we focus on topics in research computing.

In Community Studios we hold closely the belief that we are all on the same team. Even though you may be the only computing or data science practitioner in your working group, we are committed to creating an environment where all feel welcome to share their experiences. This way, despite the challenges we face in our work, we can all learn from each other and improve our skills.

```{r}
#| echo: false
#| warning: false
#| message: false
library(googlesheets4)
library(dplyr)
library(ggplot2)
library(tidyr)
library(forcats)

recode_likert <- function(raw_likert){

  levs=c(`1`="Strongly Disagree", `2`="Disagree", `3`="Neutral", `4`="Agree", `5`="Strongly Agree")
  
responses_likert <- raw_likert |>
  mutate(across(everything(), as.factor)) |>
  mutate(id=row_number()) |>
  tidyr::pivot_longer(cols=-id, names_to = "category", values_to = "likert") |>
  mutate(likert=fct_recode(likert, "Strongly Disagree"="1", "Disagree"="2", "Neutral"="3", "Agree"="4", "Strongly Agree"="5")) |>
  mutate(likert=ordered(likert)) |>
  mutate(likert = fct_rev(likert)) |>
  pivot_wider(id_cols = id, names_from=category, values_from = likert) |>
  select(-id) |>
  mutate(across(everything(), ~ factor(.x, levels = levs)))

return(responses_likert)
}
```

```{r}
#| warning: false
#| message: false

survey_data <- read_sheet(params$sheet)

```
## Community Studios Format

Community Studios is a bi-weekly meetup that covers a particular topic of interest. They differ from courses in that they focus on the **application** of computational knowledge in real-world settings by real-world people at Fred Hutch and beyond. We use Community Notes to capture discussions, questions, and share resources about topics. 

Formats for community studios include discussion of topics, show and tell presentations, short format topics, co-working sessions, and workshops. 

An atmosphere of psychological safety is the foundation of all Community Studios activities. We have [Participation Guidelines](https://hutchdatascience.org/communitystudios/guidelines) to make CommStudios welcoming and inclusive to all. This includes outlining both desired behaviors (*be collaborative*; *be respectful*) as well as behavior that is not tolerated (personal attacks, derogatory language).

## Topics Covered

The following is a list of topics discussed during the first season of RI Community Studios. Links to the Session notes are provided. 

| Date | Community Notes | Topics Discussed |
| --- | --- | --- |
| March 28, 2024 | [Session Notes](https://docs.google.com/document/d/1KzLZ9VCJW4ofaS-dXdKA8TR2ivAsVvHskJ-hQftLn-s/edit?usp=sharing) | Intro topics |
| April 4, 2024 | [Session Notes](https://docs.google.com/document/d/1Gf5VQdvyanmLLXCUg7Z3yydsQ619uuLjs3V9vzbghA8/edit?usp=sharing) | PROOF / Intro to WILDS (Taylor Firman) |
| April 18, 2024 | [Session Notes](https://docs.google.com/document/d/1xJWJV8wmO7qJd_PdAvdFWqh7cKRCyZ8IXriizV63WOQ/edit?usp=sharing)| Streamlit (Sam Minot) / 15 Tips for Success with HPC |
| May 2, 2024 | [Session Notes](https://docs.google.com/document/d/1W5Cw3iXQED4kV9fZA6L7GQUnzJzdpCW5TnN1HTN69hM/edit?usp=sharing) | Nextflow for Model Building (Diya Sashidhar) / Software Environments for Research Computing |
| May 16, 2024 | [Session Notes](https://docs.google.com/document/d/1kGbe0VAOAC4e2q7R85ZOO8je7VZIX7wPawvhPV76_Qo/edit?usp=sharing) | Research Data Management |
| May 30, 2024 | [Session Notes](https://docs.google.com/document/d/1L84N6qyk6WL9oLCqQV2b7ZTgvEZLKG7xt8rz2XoEKRs/edit?usp=sharing) | Nextflow for Isoform ID of PacBio Reads (Logan Wallace)  |

# Survey Results

We conducted a survey (n=`r nrow(survey_data)`) measuring the psychological safety and impact of community studios. Note the low n, but this is a baseline for comaprison with future seasons. 

## Psychological Safety

Almost all research informaticians and computational biologists have impostor syndrome. That's why having a forum where it's safe to ask questions is important.

Overall, RI CommStudios rated quite high in psychological safety, with the majority of members strongly agreeing with the two psychological safety measures: being a place where they feel comfortable to raise problems and concerns, and ease of approaching other community studios members for help.

```{r}
#| echo: false
responses_psych <- survey_data |> dplyr::select(contains("problems and concerns") | contains("easy to ask")) |> dplyr::rename(`I feel able\nto bring up problems\nand concerns`="On a scale of 1-5 (1-disagree, 5-agree), how much do you agree with the following statements:\n\nI feel able to bring up problems and concerns I have at community studios.")
```

```{r}
#| echo: false
responses_likert <- recode_likert(responses_psych)

responses_likert <- as.data.frame(responses_likert)

ggstats::gglikert(responses_likert, y_label_wrap = 20, labels_size = 1.5) + labs(title="CommStudios Psychological Safety is High") 
```

## Impact

All respondees agreed that RI CommStudios was useful in their day-to-day work. 70% of respondents recommended Community Studios to a Colleague.

```{r}
#| echo: false
#| fig-height: 2
responses_impact <- survey_data |> select(contains("helpful"))
responses_likert <- recode_likert(responses_impact)
ggstats::gglikert(responses_likert, y_label_wrap = 20, labels_size = 1.5)  #+ theme_barbie()
```

Attendees felt that RI CommStudios had a lot of value: 

> *I would very much like to see community studios continue and believe it has real tangible value.*

> *I really appreciated the enthusiasm and friendliness of everyone*

> *Thanks for putting these together!*

```{r}
#| eval: false
survey_data |> select(contains("recommended")) |>
  rename(`recommended to colleague` = "Have you recommended community studios to a colleague in or outside your lab?") |>  ggplot(aes(x=`recommended to colleague`, fill=`recommended to colleague`)) + geom_bar() + ggtitle("Have you recommended community studios to a colleague?")
```

## Recommendations for Future Community Studios

There is a desire for more interactive activities by participants, such as: 

- workshops  
- co-working sessions 
- code review

There is also a desire for subject specific community studios, such as working with OMICs data. 

Scheduling was the main barrier for those who could not attend. Additionally, there was feedback about making it easier to RSVP, which we will address. 

## Changes to Implement for Season 2

Given the feedback, we will want to implement the following recommendations for Season 2. 

- **Change branding to *Learning Community*.** external feedback (not shown) from other community organizers suggests that *Community Studios* is less inclusive than *learning community* - we will shift the branding over.

- **Make Zoom Meeting a single recurring invite.** We made the RSVPs to each community studios a separate invite. We recently discovered that we can make a recurring Zoom meeting for every comm studios meeting.

- **More short form content.** We experimented with shorter tips and tricks presentations, such as using the [`screen` utility in Unix](https://youtu.be/fYWszwc-LXM). We will experiment with more such presentations. 

- **Workshops.** We will experiment with workshop sessions in Research Computing topics, such as exploring containers. 

- **Have subject specific Learning Community events.** Genomics seems to be a popular topic, so we will explore bringing applications of this topic and others.

- **Increase Recruitment Efforts and Synchronize with DaSL Training**. We will increase our recruitment efforts to recruit at Fred Hutch and related institutions. 

- **Start a Learning Community [Champions Program](https://www.cscce.org/resources/community-champions-programs/).** In order to increase member engagement, we will offer a mini-mentorship program in research computing to help upskill community members.
